19Q2

### 2019-05-17 19Q2.Rmd

This comprehensive cancer cell line information was curated by both Daniel Charytonowicz and myself.
```{r dan_ccl_database}
ccl_converter <- read.delim("./data_munging/cell_line_database_v1_20180911.tsv", row.names = 1, sep = "\t", header = TRUE)
colnames(ccl_converter)[colnames(ccl_converter) == "DepMap_ID"] <- "DepMap_ID"
ccl_lineage_converter <- unique(ccl_converter[, c("DepMap_ID", "grouped_general_doid", "group_general_lineage_name")])
lineage_converter <- unique(select(ccl_lineage_converter, grouped_general_doid, group_general_lineage_name))
ccl_histologies <- read.delim("./data_munging/ccl_histologies_metadata.csv", sep = ",", header = TRUE, row.names = 1)
colnames(ccl_histologies) <- c("Cell_line", "CCLE_histology", "TCGA_classification", "CCLE_name", "DepMap_ID")
```

@@

`r nrow(filter(test, n != 1))`/`r nrow(test)` gene-cell line combinations have more than one mutation in the MAF file. This is corrected in the point mutation filtering done below.

**MERGE WITHOUT G2P AND PTMUT FILTER (5/15)**
```{r merge_filter_drug_data, eval=FALSE}
# Load in Pozdeyev et al. (2016) data
ccle_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CCLE_drug_response_metrics.csv.gz", sep = ",")
ctrp_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CTRP_drug_response_metrics.csv.gz", sep = ",")
gdsc_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/GDSC_drug_response_metrics.csv.gz", sep = ",")

@@

@@

```{r ccl_venn, eval=FALSE}
ccle_ccls <- unique(ccle$DepMap_ID)
ctrp_ccls <- unique(ctrp$DepMap_ID)
gdsc_ccls <- unique(gdsc$DepMap_ID)
# crispr_ccls <- unique(crispr_ccls$DepMap_ID)

# futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
# venn.diagram(x = list(ccle_ccls, ctrp_ccls, gdsc_ccls, crispr_ccls),
#              filename = "./plots/manuscript/all_venn_ccls.png",
#              output = TRUE, imagetype = "png",
#              height = 2000, width = 2000, resolution = 500,
#              lwd = 1, cex = 1,
#              category.names = c("CCLE", "CTRP", "GDSC", "CRISPR"),
#              cat.cex = 1.4, cat.fontface = "bold", cat.default.pos = "outer",
#              # cat.dist = c(0.06, 0.06, 0.035),
#              fill = c("springgreen4", "steelblue4", "turquoise4", "darkorchid"))
futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn.diagram(x = list(ccle_ccls, ctrp_ccls, gdsc_ccls),
             filename = "./plots/manuscript/drug_venn_ccls.png",
             output = TRUE, imagetype = "png",
             height = 2000, width = 2000, resolution = 500,
             lwd = 1, cex = 1.4,
             category.names = c("CCLE", "CTRP", "GDSC"),
             cat.cex = 1.4, cat.fontface = "bold", cat.default.pos = "outer",
             cat.dist = c(0.06, 0.06, 0.035),
             fill = c("springgreen4", "steelblue4", "turquoise4"))
```

@@

**RUN THIS WHEN CRISPR IS DONE MERGING (5/17)**
```{r histology_dists_data, eval=FALSE}
ccle_hists <- unique(select(ccle, DepMap_ID, disease)) %>% count(disease)
ccle_hists$Perc <- ccle_hists$n / sum(ccle_hists$n) * 100
ccle_hists$Dataset <- "CCLE"
ctrp_hists <- unique(select(ctrp, DepMap_ID, disease)) %>% count(disease)
ctrp_hists$Perc <- ctrp_hists$n / sum(ctrp_hists$n) * 100
ctrp_hists$Dataset <- "CTRP"
gdsc_hists <- unique(select(gdsc, DepMap_ID, disease)) %>% count(disease)
gdsc_hists$Perc <- gdsc_hists$n / sum(gdsc_hists$n) * 100
gdsc_hists$Dataset <- "GDSC"
crispr_hists <- unique(select(crispr, DepMap_ID, disease)) %>% count(disease)
crispr_hists$Perc <- crispr_hists$n / sum(crispr_hists$n) * 100
crispr_hists$Dataset <- "CRISPR"

@@

**RUN THIS WHEN CRISPR IS DONE MERGING (5/17)**
```{r histology_dists, fig.height=4.5, fig.width=4, eval=FALSE}
drug_hists <- readRDS("./data_munging/rds/histology_distributions.rds")
# png(filename = "./plots/manuscript/all_histology_distributions.png", width = 8, height = 9, units = "in", res = 500)
ggplot(drug_hists, aes(x = Dataset, y = n, fill = disease)) +
  geom_bar(stat = "identity", color = "black") +
  scale_y_continuous(breaks = seq(0, 850, by = 50), labels = seq(0, 850, by = 50)) +
  guides(fill = guide_legend(title = "Histology", ncol = 1)) +
  labs(y = "Count")
# dev.off()
```

**RUN THIS WHEN CRISPR IS DONE MERGING (5/17)**
```{r, eval=FALSE}
ccle_ccls <- unique(ccle$DepMap_ID)
ctrp_ccls <- unique(ctrp$DepMap_ID)
gdsc_ccls <- unique(gdsc$DepMap_ID)
crispr_ccls <- unique(crispr$DepMap_ID)

@@

```{r dg_heatmap, fig.height=6, fig.width=8, warning=FALSE}
g2p_dg_counts <- g2p %>% count(Drug, Gene) %>% filter(!is.na(Drug), !is.na(Gene))
g2p_dg_counts <- dcast(g2p_dg_counts, Drug ~ Gene, value.var = "n", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Drug)
g2p_dg_counts <- g2p_dg_counts[order(rowSums(-g2p_dg_counts)), order(colSums(-g2p_dg_counts))]
g2p_dg_counts[g2p_dg_counts == 0] <- NA

@@

# Define color scale
col_fun <- colorRamp2(c(0, 100), c("mistyrose", "purple4"))
# Legend properties list
ht_legend_list <- list(legend_direction = "horizontal", legend_width = unit(2.5, "in"), border = "gray70")
ht <- Heatmap(g2p_dg_counts, 
              name = "Number of G2P interpretations",
              col = col_fun,
              na_col = "gray97",
              rect_gp = gpar(col = "gray70"),
              heatmap_legend_param = ht_legend_list,
              cluster_rows = FALSE, 
              cluster_columns = FALSE, 
              show_row_names = FALSE,
              row_names_side = "left",
              row_names_gp = gpar(fontsize = 10),
              column_names_side = "bottom",
              column_names_gp = gpar(fontsize = 10),
              show_column_names = FALSE,
              bottom_annotation = HeatmapAnnotation(
                  text = anno_text(colnames(g2p_dg_counts),
                                   rot = 45,
                                   gp = gpar(fontsize = 10),
                                   offset = unit(1, "npc"),
                                   just = "right"),
                  annotation_height = unit(0.45, "in")),
              row_title = "Drugs",
              row_title_gp = gpar(fontface = "bold", fontsize = 14),
              column_title = "Genes", 
              column_title_gp = gpar(fontface = "bold", fontsize = 14),
              column_title_side = "bottom")

@@

```{r}
ccle_auc_grid <- dcast(filter(ccle_ptmuts, Drug %in% g2p_drugs), Drug ~ DepMap_ID, value.var = "Score", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Drug)
ctrp_auc_grid <- dcast(filter(ctrp_ptmuts, Drug %in% g2p_drugs), Drug ~ DepMap_ID, value.var = "Score", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Drug)
gdsc_auc_grid <- dcast(filter(gdsc_ptmuts, Drug %in% g2p_drugs), Drug ~ DepMap_ID, value.var = "Score", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Drug)

ccle_auc_ht <- Heatmap(ccle_auc_grid,
                       name = "AUC",
                       show_column_names = FALSE, 
                       column_title = "CCLE",
                       column_title_side = "top")
ctrp_auc_ht <- Heatmap(ctrp_auc_grid,
                       name = "AUC",
                       show_column_names = FALSE, 
                       column_title = "CTRP",
                       column_title_side = "top")
gdsc_auc_ht <- Heatmap(gdsc_auc_grid,
                       name = "AUC",
                       show_column_names = FALSE, 
                       column_title = "GDSC",
                       column_title_side = "top")

draw(ccle_auc_ht)
draw(ctrp_auc_ht)
draw(gdsc_auc_ht)
```

@@

```{r wilcoxon_tests, eval=FALSE}
# crispr_ptmuts <- unfactorize(crispr_ptmuts)
# # crispr_signif_all_gene <- WilcoxonTestDF(dataframe = crispr_ptmuts, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_gene.rds")
# crispr_ptmuts_filt <- filter(crispr_ptmuts, Gene %in% g2p_genes)
# crispr_signif_gene <- WilcoxonTestDF(dataframe = crispr_ptmuts_filt, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_g2p_gene.rds")
# crispr_signif_lineage <- WilcoxonTestDF(dataframe = crispr_ptmuts_filt, grouping = c("Gene", "Primary.Disease"), outfile = "./data_munging/rds/crispr_signif_g2p_lineage.rds")

ccle_filt <- filter(ccle, Drug %in% g2p_drugs & Gene %in% g2p_genes)
ccle_filt$Gene <- factor(ccle_filt$Gene, ordered = FALSE)
ccle_filt <- unfactorize(ccle_filt)
ccle_signif_dg <- WilcoxonTestDF(dataframe = ccle, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ccle_signif_dg.rds")
ccle_signif_g2p_dg <- WilcoxonTestDF(dataframe = ccle_filt, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ccle_signif_g2p_dg.rds")
ccle_signif_g2p_dgl <- WilcoxonTestDF(dataframe = ccle_filt, grouping = c("Drug", "Gene", "disease", "Drug_Gene_Lin"), outfile = "./data_munging/rds/ccle_signif_g2p_dgl.rds")

# ctrp_filt <- filter(ctrp, Drug %in% g2p_drugs)
# ctrp_filt$Gene <- factor(ctrp_filt$Gene, ordered = FALSE)
# ctrp_filt <- unfactorize(ctrp_filt)
# ctrp_signif_gene <- WilcoxonTestDF(dataframe = ctrp, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ctrp_signif_g2p_gene.rds")
# ctrp_signif_g2p_druggene <- WilcoxonTestDF(dataframe = filter(ctrp, Drug %in% g2p_drugs & Gene %in% g2p_genes), grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ctrp_signif_g2p_druggene.rds")
# ctrp_signif_lineage <- WilcoxonTestDF(dataframe = ctrp_filt, grouping = c("Drug", "Gene", "Primary.Disease", "Drug_Gene_Lin"), outfile = "./data_munging/rds/ctrp_signif_g2p_lineage.rds")
# 
# gdsc_filt <- filter(gdsc, Drug %in% g2p_drugs)
# gdsc_filt$Gene <- factor(gdsc_filt$Gene, ordered = FALSE)
# gdsc_filt <- unfactorize(gdsc_filt)
# gdsc_signif_gene <- WilcoxonTestDF(dataframe = gdsc, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/gdsc_signif_g2p_gene.rds")
# gdsc_signsif_g2p_druggene <- WilcoxonTestDF(dataframe = filter(gdsc, Drug %in% g2p_drugs & Gene %in% g2p_genes), grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/gdsc_signif_g2p_druggene.rds")
# gdsc_signif_lineage <- WilcoxonTestDF(dataframe = gdsc_filt, grouping = c("Drug", "Gene", "Primary.Disease", "Drug_Gene_Lin"), outfile = "./data_munging/rds/gdsc_signif_g2p_lineage.rds")
```

@@

ccle_signif_gene <- readRDS("./data_munging/rds/ccle_signif_g2p_gene.rds")
ccle_signif_g2p_druggene <- readRDS("./data_munging/rds/ccle_signif_g2p_druggene.rds")
ccle_signif_lineage <- readRDS("./data_munging/rds/ccle_signif_g2p_lineage.rds")

ctrp_signif_gene <- readRDS("./data_munging/rds/ctrp_signif_g2p_gene.rds")
ctrp_signif_g2p_druggene <- readRDS("./data_munging/rds/ctrp_signif_g2p_druggene.rds")
ctrp_signif_lineage <- readRDS("./data_munging/rds/ctrp_signif_g2p_lineage.rds")

gdsc_signif_gene <- readRDS("./data_munging/rds/gdsc_signif_g2p_gene.rds")
gdsc_signif_g2p_druggene <- readRDS("./data_munging/rds/gdsc_signif_g2p_druggene.rds")
gdsc_signif_lineage <- readRDS("./data_munging/rds/gdsc_signif_g2p_lineage.rds")
```

>>>>

This comprehensive cancer cell line information was curated by both Daniel Charytonowicz and myself.
```{r dan_ccl_database}
ccl_converter <- read.delim("./data_munging/cell_line_database_v1_20180911.tsv", row.names = 1, sep = "\t", header = TRUE)
colnames(ccl_converter)[colnames(ccl_converter) == "Broad_ID"] <- "DepMap_ID"
ccl_lineage_converter <- unique(ccl_converter[, c("DepMap_ID", "grouped_general_doid", "group_general_lineage_name")])
lineage_converter <- unique(select(ccl_lineage_converter, grouped_general_doid, group_general_lineage_name))
ccl_histologies <- read.delim("./data_munging/ccl_histologies_metadata.csv", sep = ",", header = TRUE, row.names = 1)
colnames(ccl_histologies) <- c("Cell_line", "CCLE_histology", "TCGA_classification", "CCLE_name", "DepMap_ID")
```

@@

`r nrow(filter(test, n != 1))`/`r nrow(test)` gene-cell line combinations have more than one mutation in the MAF file. This is corrected in the point mutation filtering done below.
```{r merge_drug_data, eval=FALSE}
# Load in Pozdeyev et al. (2016) data
ccle_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CCLE_drug_response_metrics.csv.gz", sep = ",")
ctrp_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CTRP_drug_response_metrics.csv.gz", sep = ",")
gdsc_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/GDSC_drug_response_metrics.csv.gz", sep = ",")

@@

## CRISPR

```{r merge_crispr_data, eval=TRUE}
crispr_all <- read.delim("./data_munging/19Q2/Achilles_gene_effect_20190514.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr_all) <- gsub(" .*", "", colnames(crispr_all))
colnames(crispr_all)[1] <- "DepMap_ID"

# Melt CRISPR dataset for merging
# crispr_melt <- melt(crispr_all, id.vars = "DepMap_ID", measure.vars = colnames(crispr_all)[2:ncol(crispr_all)], variable.name = "Gene", value.name = "Score")

crispr <- MergeOmicsG2P(database = "CRISPR", maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = g2p_genes, outfile = "./data_munging/19Q2/crispr_data_19Q2.rds")
write.csv.gz(crispr %>% drop_na(Score), "./data_munging/19Q2/crispr_data_19Q2.csv.gz")
```

@@

```{r ccl_venn, eval=TRUE}
ccle_ccls <- unique(ccle$DepMap_ID)
ctrp_ccls <- unique(ctrp$DepMap_ID)
gdsc_ccls <- unique(gdsc$DepMap_ID)
crispr_ccls <- unique(crispr_ccls$DepMap_ID)

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn.diagram(x = list(ccle_ccls, ctrp_ccls, gdsc_ccls, crispr_ccls),
             filename = "./plots/manuscript/all_venn_ccls.png",
             output = TRUE, imagetype = "png",
             height = 2000, width = 2000, resolution = 500,
             lwd = 1, cex = 1,
             category.names = c("CCLE", "CTRP", "GDSC", "CRISPR"),
             cat.cex = 1.4, cat.fontface = "bold", cat.default.pos = "outer",
             # cat.dist = c(0.06, 0.06, 0.035),
             fill = c("springgreen4", "steelblue4", "turquoise4", "darkorchid"))
futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn.diagram(x = list(ccle_ccls, ctrp_ccls, gdsc_ccls),
             filename = "./plots/manuscript/drug_venn_ccls.png",
             output = TRUE, imagetype = "png",
             height = 2000, width = 2000, resolution = 500,
             lwd = 1, cex = 1.4,
             category.names = c("CCLE", "CTRP", "GDSC"),
             cat.cex = 1.4, cat.fontface = "bold", cat.default.pos = "outer",
             cat.dist = c(0.06, 0.06, 0.035),
             fill = c("springgreen4", "steelblue4", "turquoise4"))
```

@@

**RUN THIS WHEN CRISPR IS DONE MERGING (5/17)**
```{r histology_dists_data, eval=TRUE}
ccle_hists <- unique(select(ccle, DepMap_ID, disease)) %>% count(disease)
ccle_hists$Perc <- ccle_hists$n / sum(ccle_hists$n) * 100
ccle_hists$Dataset <- "CCLE"
ctrp_hists <- unique(select(ctrp, DepMap_ID, disease)) %>% count(disease)
ctrp_hists$Perc <- ctrp_hists$n / sum(ctrp_hists$n) * 100
ctrp_hists$Dataset <- "CTRP"
gdsc_hists <- unique(select(gdsc, DepMap_ID, disease)) %>% count(disease)
gdsc_hists$Perc <- gdsc_hists$n / sum(gdsc_hists$n) * 100
gdsc_hists$Dataset <- "GDSC"
crispr_hists <- unique(select(crispr, DepMap_ID, disease)) %>% count(disease)
crispr_hists$Perc <- crispr_hists$n / sum(crispr_hists$n) * 100
crispr_hists$Dataset <- "CRISPR"

@@

**RUN THIS WHEN CRISPR IS DONE MERGING (5/17)**
```{r histology_dists, fig.height=4.5, fig.width=4, eval=TRUE}
drug_hists <- readRDS("./data_munging/rds/histology_distributions.rds")
png(filename = "./plots/manuscript/all_histology_distributions.png", width = 8, height = 9, units = "in", res = 500)
ggplot(drug_hists, aes(x = Dataset, y = n, fill = disease)) +
  geom_bar(stat = "identity", color = "black") +
  scale_y_continuous(breaks = seq(0, 850, by = 50), labels = seq(0, 850, by = 50)) +
  guides(fill = guide_legend(title = "Histology", ncol = 1)) +
  labs(y = "Count")
dev.off()
```

**RUN THIS WHEN CRISPR IS DONE MERGING (5/17)**
```{r, eval=TRUE}
ccle_ccls <- unique(ccle$DepMap_ID)
ctrp_ccls <- unique(ctrp$DepMap_ID)
gdsc_ccls <- unique(gdsc$DepMap_ID)
crispr_ccls <- unique(crispr$DepMap_ID)

@@

```{r dg_heatmap, fig.height=6, fig.width=8, warning=FALSE}
g2p_dg_counts <- g2p %>% count(Drug, Gene) %>% filter(!is.na(Drug), !is.na(Gene))
g2p_dg_counts <- dcast(g2p_dg_counts, Drug ~ Gene, value.var = "n", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Drug)
g2p_dg_counts <- g2p_dg_counts[order(rowSums(-g2p_dg_counts)), order(colSums(-g2p_dg_counts))]
# g2p_dg_counts[g2p_dg_counts == 0] <- NA

@@

# Define color scale
col_fun <- colorRamp2(c(0, 1, 100), c("gray97", "mistyrose", "purple4"))
# Legend properties list
ht_legend_list <- list(legend_direction = "horizontal", legend_width = unit(2.5, "in"), border = "gray70")
ht <- Heatmap(g2p_dg_counts, 
              name = "Number of G2P interpretations",
              col = col_fun,
              na_col = "gray97",
              rect_gp = gpar(col = "gray70"),
              heatmap_legend_param = ht_legend_list,
              cluster_rows = FALSE, 
              cluster_columns = FALSE, 
              show_row_names = FALSE,
              row_names_side = "left",
              row_names_gp = gpar(fontsize = 10),
              column_names_side = "bottom",
              column_names_gp = gpar(fontsize = 10),
              show_column_names = FALSE,
              bottom_annotation = HeatmapAnnotation(
                  text = anno_text(colnames(g2p_dg_counts),
                                   rot = 45,
                                   gp = gpar(fontsize = 10),
                                   offset = unit(1, "npc"),
                                   just = "right"),
                  annotation_height = unit(0.45, "in")),
              row_title = "Drugs",
              row_title_gp = gpar(fontface = "bold", fontsize = 14),
              column_title = "Genes", 
              column_title_gp = gpar(fontface = "bold", fontsize = 14),
              column_title_side = "bottom")

@@

```{r auc_clustered_ht, eval=FALSE}
ccle_auc_grid <- dcast(ccle, Drug ~ DepMap_ID, value.var = "Score", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Drug)
ctrp_auc_grid <- dcast(ctrp, Drug ~ DepMap_ID, value.var = "Score", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Drug)
gdsc_auc_grid <- dcast(gdsc, Drug ~ DepMap_ID, value.var = "Score", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Drug)

ccle_lineages_ha <- merge(data.frame("DepMap_ID" = colnames(ccle_auc_grid)), select(ccl_meta, DepMap_ID, disease), by = "DepMap_ID", all.x = TRUE) %>% `rownames<-`(.[,1]) %>% select(-DepMap_ID)
ctrp_lineages_ha <- merge(data.frame("DepMap_ID" = colnames(ctrp_auc_grid)), select(ccl_meta, DepMap_ID, disease), by = "DepMap_ID", all.x = TRUE) %>% `rownames<-`(.[,1]) %>% select(-DepMap_ID)
gdsc_lineages_ha <- merge(data.frame("DepMap_ID" = colnames(gdsc_auc_grid)), select(ccl_meta, DepMap_ID, disease), by = "DepMap_ID", all.x = TRUE) %>% `rownames<-`(.[,1]) %>% select(-DepMap_ID)

ht_legend_list <- list(direction = "horizontal")

ccle_col_fun <- colorRamp2(c(0, 1), c("gray97", "darkgreen"))
ccle_auc_ha <- HeatmapAnnotation(Lineage = ccle_lineages_ha$disease, annotation_legend_param = list(Lineage = list(direction = "horizontal", nrow = 5)))
ccle_auc_ht <- Heatmap(ccle_auc_grid,
                       na_col = "black",
                       col = ccle_col_fun,
                       top_annotation = ccle_auc_ha,
                       heatmap_legend_param = ht_legend_list,
                       name = "AUC",
                       show_column_names = FALSE, 
                       row_title = "Drugs and Small Molecules",
                       column_title = "Cancer Cell Lines",
                       column_title_side = "top")

ctrp_col_fun <- colorRamp2(c(0, 1), c("gray97", "steelblue4"))
ctrp_auc_ha <- HeatmapAnnotation(Lineage = ctrp_lineages_ha$disease, annotation_legend_param = list(Lineage = list(direction = "horizontal", nrow = 3)))
ctrp_auc_ht <- Heatmap(ctrp_auc_grid,
                       na_col = "black",
                       col = ctrp_col_fun,
                       top_annotation = ctrp_auc_ha,
                       heatmap_legend_param = ht_legend_list,
                       name = "AUC",
                       show_column_names = FALSE, 
                       row_title = "Drugs and Small Molecules",
                       column_title = "Cancer Cell Lines",
                       column_title_side = "top")

gdsc_col_fun <- colorRamp2(c(0, 1), c("gray97", "turquoise4"))
gdsc_auc_ha <- HeatmapAnnotation(Lineage = gdsc_lineages_ha$disease, annotation_legend_param = list(Lineage = list(direction = "horizontal", nrow = 3)))
gdsc_auc_ht <- Heatmap(gdsc_auc_grid,
                       na_col = "black",
                       col = gdsc_col_fun,
                       top_annotation = gdsc_auc_ha,
                       heatmap_legend_param = ht_legend_list,
                       name = "AUC",
                       show_column_names = FALSE, 
                       row_title = "Drugs and Small Molecules",
                       column_title = "Cancer Cell Lines",
                       column_title_side = "top")

png(filename = "./plots/manuscript/ccle_clustered_heatmap.png", width = 10, height = 8, units = "in", res = 500)
draw(ccle_auc_ht, merge_legend = TRUE, heatmap_legend_side = "top", annotation_legend_side = "top")
dev.off()

png(filename = "./plots/manuscript/ctrp_clustered_heatmap.png", width = 22, height = 40, units = "in", res = 500)
draw(ctrp_auc_ht, merge_legend = TRUE, heatmap_legend_side = "top", annotation_legend_side = "top")
dev.off()

png(filename = "./plots/manuscript/gdsc_clustered_heatmap.png", width = 18, height = 20, units = "in", res = 500)
draw(gdsc_auc_ht, merge_legend = TRUE, heatmap_legend_side = "top", annotation_legend_side = "top")
dev.off()
```

@@

```{r wilcoxon_tests, eval=TRUE}
# crispr_ptmuts <- unfactorize(crispr_ptmuts)
# # crispr_signif_all_gene <- WilcoxonTestDF(dataframe = crispr_ptmuts, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_gene.rds")
# crispr_ptmuts_filt <- filter(crispr_ptmuts, Gene %in% g2p_genes)
# crispr_signif_gene <- WilcoxonTestDF(dataframe = crispr_ptmuts_filt, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_g2p_gene.rds")
# crispr_signif_lineage <- WilcoxonTestDF(dataframe = crispr_ptmuts_filt, grouping = c("Gene", "Primary.Disease"), outfile = "./data_munging/rds/crispr_signif_g2p_lineage.rds")

# ccle_filt <- filter(ccle, Drug %in% g2p_drugs & Gene %in% g2p_genes)
# ccle_filt$Gene <- factor(ccle_filt$Gene, ordered = FALSE)
# ccle_filt <- unfactorize(ccle_filt)
# ccle_signif_dg <- WilcoxonTestDF(dataframe = ccle, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ccle_signif_dg.rds")
# ccle_signif_g2p_dg <- WilcoxonTestDF(dataframe = ccle_filt, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ccle_signif_g2p_dg.rds")
# ccle_signif_g2p_dgl <- WilcoxonTestDF(dataframe = ccle_filt, grouping = c("Drug", "Gene", "disease", "Drug_Gene_Lin"), outfile = "./data_munging/rds/ccle_signif_g2p_dgl.rds")

ctrp_filt <- filter(ctrp, Drug %in% g2p_drugs & Gene %in% g2p_genes)
ctrp_filt$Gene <- factor(ctrp_filt$Gene, ordered = FALSE)
ctrp_filt <- unfactorize(ctrp_filt)
ctrp_signif_dg <- WilcoxonTestDF(dataframe = ctrp, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ctrp_signif_dg.rds")
ctrp_signif_g2p_dg <- WilcoxonTestDF(dataframe = ctrp_filt, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ctrp_signif_g2p_dg.rds")
ctrp_signif_g2p_dgl <- WilcoxonTestDF(dataframe = ctrp_filt, grouping = c("Drug", "Gene", "disease", "Drug_Gene_Lin"), outfile = "./data_munging/rds/ctrp_signif_g2p_dgl.rds")

gdsc_filt <- filter(gdsc, Drug %in% g2p_drugs & Gene %in% g2p_genes)
gdsc_filt$Gene <- factor(gdsc_filt$Gene, ordered = FALSE)
gdsc_filt <- unfactorize(gdsc_filt)
gdsc_signif_dg <- WilcoxonTestDF(dataframe = gdsc, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/gdsc_signif_dg.rds")
gdsc_signif_g2p_dg <- WilcoxonTestDF(dataframe = gdsc_filt, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/gdsc_signif_g2p_dg.rds")
gdsc_signif_g2p_dgl <- WilcoxonTestDF(dataframe = gdsc_filt, grouping = c("Drug", "Gene", "disease", "Drug_Gene_Lin"), outfile = "./data_munging/rds/gdsc_signif_g2p_dgl.rds")
```

@@

# ccle_signif_gene <- readRDS("./data_munging/rds/ccle_signif_g2p_gene.rds")
# ccle_signif_g2p_druggene <- readRDS("./data_munging/rds/ccle_signif_g2p_druggene.rds")
# ccle_signif_lineage <- readRDS("./data_munging/rds/ccle_signif_g2p_lineage.rds")
# 
# ctrp_signif_gene <- readRDS("./data_munging/rds/ctrp_signif_g2p_gene.rds")
# ctrp_signif_g2p_druggene <- readRDS("./data_munging/rds/ctrp_signif_g2p_druggene.rds")
# ctrp_signif_lineage <- readRDS("./data_munging/rds/ctrp_signif_g2p_lineage.rds")
# 
# gdsc_signif_gene <- readRDS("./data_munging/rds/gdsc_signif_g2p_gene.rds")
# gdsc_signif_g2p_druggene <- readRDS("./data_munging/rds/gdsc_signif_g2p_druggene.rds")
# gdsc_signif_lineage <- readRDS("./data_munging/rds/gdsc_signif_g2p_lineage.rds")
```

# SPEARMAN (CN, GE)

```{r drug_spear_calc, warning=FALSE, eval=TRUE}
g2p_druggene_tally <- g2p %>% group_by(Drug_Gene) %>% tally() %>% drop_na()

# CCLE
ccle_cor_res <- ccle %>% group_by(Drug, Gene, Drug_Gene) %>% summarize(
                GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate,
                GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value,
                CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
                CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value)
ccle_cor_res <- merge(g2p_druggene_tally, ccle_cor_res, by = "Drug_Gene", all.y = TRUE)
ccle_cor_res$n <- ifelse(is.na(ccle_cor_res$n), 0, ccle_cor_res$n)
ccle_cor_res$InG2P <- ifelse(ccle_cor_res$n != 0, "Yes", "No")
ccle_cor_res$Dataset <- "CCLE"

# CTRP
ctrp_cor_res <- ctrp %>% group_by(Drug, Gene, Drug_Gene) %>% summarize(
                GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate,
                GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value,
                CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
                CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value)
ctrp_cor_res <- merge(g2p_druggene_tally, ctrp_cor_res, by = "Drug_Gene", all.y = TRUE)
ctrp_cor_res$n <- ifelse(is.na(ctrp_cor_res$n), 0, ctrp_cor_res$n)
ctrp_cor_res$InG2P <- ifelse(ctrp_cor_res$n != 0, "Yes", "No")
ctrp_cor_res$Dataset <- "CTRP"

# GDSC
gdsc_cor_res <- gdsc %>% group_by(Drug, Gene, Drug_Gene) %>% summarize(
                GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate,
                GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value,
                CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
                CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value)
gdsc_cor_res <- merge(g2p_druggene_tally, gdsc_cor_res, by = "Drug_Gene", all.y = TRUE)
gdsc_cor_res$n <- ifelse(is.na(gdsc_cor_res$n), 0, gdsc_cor_res$n)
gdsc_cor_res$InG2P <- ifelse(gdsc_cor_res$n != 0, "Yes", "No")
gdsc_cor_res$Dataset <- "GDSC"

# All
g2p_cor <- rbind(ccle_cor_res, ctrp_cor_res, gdsc_cor_res)

saveRDS(g2p_cor, "./data_munging/rds/drug_cor.rds")
```

```{r load_drug_spear}
g2p_cor <- readRDS("./data_munging/rds/drug_cor.rds")
```

Spearman correlations for gene expression and copy number:
```{r crispr_spear_calc, warning=FALSE}
crispr_spear_ge <- transform(crispr_ptmuts_filt %>% group_by(Gene) %>% summarize(p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value, p_log10 = -log10(p), corr = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate, Metric = "Gene Expression"), Gene = reorder(Gene, corr))
crispr_spear_cn <- transform(crispr_ptmuts_filt %>% group_by(Gene) %>% summarize(p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value, p_log10 = -log10(p), corr = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate, Metric = "Copy Number"), Gene = reorder(Gene, corr))
```
